---
title: "Learn FindVariableFeatures"
author: "Juechen Yang"
date: "5/18/2022"
output: rmarkdown::github_document
---

## load libraries
```{r warning=FALSE, message=FALSE}
library(Seurat)
library(SeuratData)
library(patchwork)
#InstallData("ifnb")
LoadData("ifnb")
```

## create object data and Normalize it
```{r}
ifnb.list <- SplitObject(ifnb, split.by = "stim")
object = ifnb.list[[1]]
#normalization
object = NormalizeData(object = object, verbose = F)
```

## get assay data for object

Because FindVariableFeatures needs the count matrix for input, so it is required to fetch the assay data from the Seurat object for downstream calculation

```{r}
object = GetAssayData(object)
#initializa verbose
verbose = FALSE
clip.max = "auto"
if (clip.max == "auto") {
      clip.max <- sqrt(x = ncol(x = object))
}
```

## computes the mean and variance for each row(gene)

```{r warning=FALSE, message=FALSE}
SparseRowVar2 <- function(mat, mu, display_progress) {
    .Call('_Seurat_SparseRowVar2', PACKAGE = 'Seurat', mat, mu, display_progress)
}
hvf.info <- data.frame(mean = rowMeans(x = object))
hvf.info$variance <- SparseRowVar2(mat = object, mu = hvf.info$mean, 
                      display_progress = verbose)
hvf.info[1:10,]
```

## assign 0 to some metrics (I don't know why doing this)

```{r}
hvf.info$variance.expected <- 0
hvf.info$variance.standardized <- 0
not.const <- hvf.info$variance > 0
```

## produce a fit using loess

This is try to regress and smooth the curve of log10(mean) and log10(var). Loess Regression is the most common method used to smooth a volatile time series. It is a non-parametric methods where least squares regression is performed in localized subsets, which makes it a suitable candidate for smoothing any numerical vector.

```{r warning=FALSE, message=FALSE}
loess.span = 0.3
fit <- loess(formula = log10(x = variance) ~ log10(x = mean), 
      data = hvf.info[not.const, ], span = loess.span)
```

## get the expected variance for gene that has larger than 0 variance

```{r warning=FALSE, message=FALSE}
hvf.info$variance.expected[not.const] <- 10^fit$fitted
```

## calculate variance of zscore which is captured by mean and expected variance

```{r}
SparseRowVarStd <- function(mat, mu, sd, vmax, display_progress) {
    .Call('_Seurat_SparseRowVarStd', PACKAGE = 'Seurat', mat, mu, sd, vmax, display_progress)
}
hvf.info$variance.standardized <- SparseRowVarStd(mat = object, 
      mu = hvf.info$mean, sd = sqrt(hvf.info$variance.expected), 
      vmax = clip.max, display_progress = verbose)
```

## we selected top n genes that have highest standardized variance
```{r}
hvf.info = hvf.info[order(hvf.info$variance.standardized, decreasing = T),]
top_n = 2000
top.genes = rownames(hvf.info)[1:top_n]
top.genes[1:20]
```
