---
title: "Learn SelectIntegrationFeatures"
author: "Juechen Yang"
date: "5/18/2022"
output: rmarkdown::github_document
---

## load libraries
```{r warning=FALSE, message=FALSE}
library(Seurat)
library(SeuratData)
library(patchwork)
#InstallData("ifnb")
LoadData("ifnb")
```


## preprocess
```{r}
ifnb.list <- SplitObject(ifnb, split.by = "stim")
ifnb.list <- lapply(X = ifnb.list, FUN = function(x) {
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})
```

## initializa parameters and preprocess
```{r warning=FALSE}

nfeatures = 2000
assay = NULL
verbose = TRUE
fvf.nfeatures = 2000
object.list = ifnb.list
```

## assign assays to each obj in the list

```{r}
if (!is.null(x = assay)) {
    if (length(x = assay) != length(x = object.list)) {
      stop("If specifying the assay, please specify one assay per object in the object.list")
    }
    for (ii in length(x = object.list)) {
      DefaultAssay(object = object.list[[ii]]) <- assay[ii]
    }
}else {
  assay <- sapply(X = object.list, FUN = DefaultAssay)
}
```

## check if variable features exist

```{r}
for (ii in 1:length(x = object.list)) {
  if (length(x = VariableFeatures(object = object.list[[ii]])) == 
    0) {
    if (verbose) {
      message(paste0("No variable features found for object", 
        ii, " in the object.list. Running FindVariableFeatures ..."))
    }
    object.list[[ii]] <- FindVariableFeatures(object = object.list[[ii]], 
      nfeatures = fvf.nfeatures, verbose = verbose, 
      ...)
  }
}
```

## fetch highly variable feature names from all seurat obj in the list

```{r}
var.features <- unname(obj = unlist(x = lapply(X = 1:length(x = object.list), 
    FUN = function(x) VariableFeatures(object = object.list[[x]], 
      assay = assay[x]))))
```

## find shared highly variable genes

```{r}
var.features <- sort(x = table(var.features), decreasing = TRUE)
#make sure highly variable genes should present in all seurat obj 
#(even may not highly variable in every)
for (i in 1:length(x = object.list)) {
  var.features <- var.features[names(x = var.features) %in% 
    rownames(x = object.list[[i]][[assay[i]]])]
}
```

## select features that have highly variable freq larger than tie value

```{r}
tie.val <- var.features[min(nfeatures, length(x = var.features))]
features <- names(x = var.features[which(x = var.features > 
    tie.val)])
length(features)
features[1:20]
```

## fetch variable features for each dataset

```{r}
vf.list <- lapply(X = object.list, FUN = VariableFeatures)
vf.list[[1]][1:10]
```

## re-rank features based on their median rank for all vairable gene set from each dataset 

```{r}
#features are genes that shared by more than tie.val datasets
if (length(x = features) > 0) {
  feature.ranks <- sapply(X = features, FUN = function(x) {
    ranks <- sapply(X = vf.list, FUN = function(vf) {
      if (x %in% vf) {
        return(which(x = x == vf))
      }
      return(NULL)
    })
    median(x = unlist(x = ranks))
  })
  features <- names(x = sort(x = feature.ranks))
}
```

## grab the features that equal to tie.val

```{r}
features.tie <- var.features[which(x = var.features == tie.val)]
```

## do the same re-rank for tie features

```{r}
tie.ranks <- sapply(X = names(x = features.tie), FUN = function(x) {
  ranks <- sapply(X = vf.list, FUN = function(vf) {
    if (x %in% vf) {
      return(which(x = x == vf))
    }
    return(NULL)
  })
  median(x = unlist(x = ranks))
})
```

## return a vector with features in the front and tie.features in the back

```{r}
features <- c(features, names(x = head(x = sort(x = tie.ranks), 
    nfeatures - length(x = features))))
```